\documentclass[two column]{article}
\usepackage{graphicx} 
\usepackage[letterpaper,top=2cm,bottom=2cm,left=3cm,right=3cm,marginparwidth=1cm]{geometry}
\usepackage[colorlinks=true, allcolors=blue]{hyperref}

\title{Smart Wearable System for Pediatric Asthma Monitoring}
\author{Bozhena Bursalı, Selin Atakan}
\date{April 2026}

\begin{document}
\maketitle
\begin{abstract}
Managing pediatric asthma is difficult because most current tools are reactive, only acting after symptoms appear. This paper introduces an IoT-based wearable system that monitors both a child's health (pulse and $SpO_2$) and their environment (air quality and humidity) in real-time. Using an ESP32-based prototype and data fusion, the system aims to predict asthma attacks before they happen. Our goal is to provide a low-cost, proactive solution that helps families manage asthma more effectively and reduces emergency room visits.
\end{abstract}

\section{Introduction}
Asthma is a major health issue for children, often triggered by environmental factors like air pollution and humidity\cite{Pubmed2026}. Currently, most families manage asthma "reactively," meaning they only use medicine after a child starts struggling to breathe. This increases the risk of serious attacks and hospital visits.

While there are many asthma apps available, most lack real-time data. For example, apps like AsthmaMD require manual input, and systems like Propeller Health do not track heart rate or oxygen levels. There is a clear need for a tool that combines body sensors with environmental data.

In this project, I develope a smart wearable prototype using an ESP32 microcontroller\cite{Li2026}. This system collects physiological data (pulse, $SpO_2$) and environmental data (PM2.5, temperature) to provide an early warning for potential attacks \cite{Pubmed2026}. By using "data fusion," we can offer a more accurate and proactive way to keep children safe.\cite{WHO}

\subsection{Regional Context and the Reactive Gap}
In Turkey and the surrounding regions, the challenges of pediatric asthma are amplified by rapid urbanization and environmental surge events. Recent data from Istanbul indicates a 9\% year-over-year increase in air pollution, with over 90\% of the urban population exposed to air quality that adversely impacts respiratory health. In such overwhelming healthcare systems that lack proactive warning mechanisms.\cite{IstanbulPollution2026}

Furthermore, the economic burden of reactive care is significant. In Turkey, the cost of managing uncontrolled asthma is nearly double that of well-controlled cases, primarily due to preventable hospitalizations. By integrating real-time environmental triggers like PM2.5 and humidity—which show the strongest correlation with pediatric ER visits in the Middle East,\cite{SaudiAsthma2025} this project seeks to bridge the gap between evidence-based recommendations and the current reality of reactive pediatric care.






\bibliographystyle{IEEEtran}
\bibliography{references}

\end{document}